Figures for Paper

### 2021-09-10 paper_figures.Rmd
more figs

---
title: "Figures for Paper"
author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins"
date: "06/10/2020"
output:
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'docs/index.html')) })
---

@@

nmd_plot = ggbarplot(for_bar_plot_sy, 
                     x = 'Gene', 
                     add = c("mean_se","jitter"), 
                     y = 'Value',
                     color = 'Condition', 
                     position = position_dodge(0.8), 
                     dot.size = 10) +
  scale_y_continuous() + 
  ggpubr::theme_pubr() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") +
  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene")
plot(nmd_plot)

@@

@@

cell_lines = fread(file.path(here::here(),"data","unc13b_salmon_cell_lines.csv"))
cell_lines %>% 
    mutate(sample = gsub(".Aligned.sorted.out","",sample)) %>% 
    left_join(rel_rna_cryptic_amount[,.(condition,plot_name,sample)]) %>% 
    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
    filter(Name.y %in% c("UNC13B-207",
                         "UNC13B-208",
                         "UNC13B-210")) %>%
    ggplot(aes(x =condition, y = TPM,fill = Name.y)) + 
    geom_boxplot() +
   geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
    # facet_wrap(~plot_name) +
  ggpubr::theme_pubr() + 
  labs(fill="",x="") + 
  scale_fill_manual(values = iso_colors)

@@

unc13b_expression = fread(file.path(here::here(),"data","unc13b_transcript_counts.csv"))

clean_data_table %>% 
    left_join(unc13b_expression) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    select(sample,disease_group2,tissue_clean,unc13a_detected,TPM,Name) %>% 
    filter(!is.na(Name)) %>% 
    unique() %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
        filter(Name.y %in% c("UNC13B-207",
                         "UNC13B-208",
                         "UNC13B-210")) %>%
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2, y = TPM, fill = Name.y)) + 
    geom_boxplot() +
    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
    facet_wrap(~tissue_clean,scales = 'free') + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) + 
    scale_y_continuous(label=comma) +
    ggpubr::theme_pubr() + 
    labs(fill="",x="") +
    scale_fill_manual(values = iso_colors)

@@

@@

>>>>

---
title: "Figures for Paper"
author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins + Matthew Keuss"
date: "06/10/2020"
output:
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'docs/index.html')) })
---

@@

nmd_plot = ggbarplot(for_bar_plot_sy, 
                     x = 'Gene', 
                     add = c("mean_se","jitter"), 
                     y = 'Value',
                     color = 'Condition',
                     fill = "Condition",
                     position = position_dodge(0.8), 
                     dot.size = 20,
                     size = 1.5) +
  scale_y_continuous() + 
  ggpubr::theme_pubr() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") +
  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene") +
    theme(axis.ticks.length=unit(0.1,"inch"),
        axis.line = element_line(colour = 'black', size = 1.5),
        axis.ticks = element_line(colour = "black", size = 2))+
  scale_fill_manual(values = c("#fca361ff","#aac043ff")) +
  scale_color_manual(values = c("#000000ff","#000000ff"))

plot(nmd_plot)

@@

## UNC13B fsE across NYGC cohort

```{r do_stats}

for_stats = clean_data_table %>% 
    filter(tissue_clean %in%
               c("Frontal_Cortex",
                 "Lumbar_Spinal_Cord",
                 "Cervical_Spinal_Cord",
                 "Motor_Cortex",
                 "Temporal_Cortex",
                 "Cerebellum") )%>%
    group_by(disease_group2,tissue_clean) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    ungroup() %>%
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
    mutate(tissue_clean = fct_relevel(tissue_clean, "Motor_Cortex",
                                      "Frontal_Cortex",
                                      "Temporal_Cortex",
                                      "Cervical_Spinal_Cord",
                                      "Lumbar_Spinal_Cord")) %>% 
  mutate(all_the_fse_reads =  (((UNC13B_nmd_junction_5prime + UNC13B_nmd_junction_3prime)/ library_size) * 10^6)) %>% 
  dplyr::select(tissue_clean,all_the_fse_reads,disease_group2) %>% 
  group_by(tissue_clean) %>% 
  nest()

stars_table = for_stats %>% 
  mutate(wilcox_test = map(data, ~{pairwise.wilcox.test(.x$all_the_fse_reads,.x$disease_group2) %>% broom::tidy()})) %>% 
  dplyr::select(-data) %>% 
  unnest() %>% 
  ungroup() %>% 
  mutate(stars = gtools::stars.pval(p.value = p.value)) %>% 
  filter(p.value < 0.05) %>% 
  mutate(y.position = seq(from = 1.5,to = 2.6,length.out = 11))

```{r plot_unc13b_fsE_nygc}
clean_data_table %>% 
    filter(tissue_clean %in%
               c("Frontal_Cortex",
                 "Lumbar_Spinal_Cord",
                 "Cervical_Spinal_Cord",
                 "Motor_Cortex",
                 "Temporal_Cortex",
                 "Cerebellum") )%>%
    group_by(disease_group2,tissue_clean) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    ungroup() %>%
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
    mutate(tissue_clean = fct_relevel(tissue_clean, "Motor_Cortex",
                                      "Frontal_Cortex",
                                      "Temporal_Cortex",
                                      "Cervical_Spinal_Cord",
                                      "Lumbar_Spinal_Cord")) %>% 
  mutate(all_the_fse_reads =  (((UNC13B_nmd_junction_5prime + UNC13B_nmd_junction_3prime)/ library_size) * 10^6)) %>% 
  ggplot(aes(x = disease_group2,
               y =all_the_fse_reads,
               fill ="red")) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) +
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
    ggplot2::facet_wrap(~tissue_clean,scales = "free",nrow = 2) +
    ylab("UNC13B fsE \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") + 
  stat_pvalue_manual(stars_table, label = "{stars}")

@@

cell_lines = fread(file.path(here::here(),"data","unc13b_salmon_cell_lines.csv"))
cell_lines %>% 
    mutate(sample = gsub(".Aligned.sorted.out","",sample)) %>% 
    left_join(rel_rna_cryptic_amount[,.(condition,plot_name,sample)]) %>% 
    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
    filter(Name.y %in% c("UNC13B-207",
                         "UNC13B-208",
                         "UNC13B-210")) %>%
    ggplot(aes(x =condition, y = TPM,fill = Name.y),color = 'black') + 
    geom_boxplot(size = 1.2,color = "black") +
   geom_point(stroke = 1.2,size = 2.3,pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
    # facet_wrap(~plot_name) +
  ggpubr::theme_pubr() + 
  labs(fill="",x="") + 
  scale_fill_manual(values = iso_colors)

@@

unc13b_expression = fread(file.path(here::here(),"data","unc13b_transcript_counts.csv"))
unc13b_expression %>% 
    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
    filter(sample != "") %>% 
    dplyr::select(sample,
                  EffectiveLength,
                  NumReads,
                  Name.y) %>% 
    unique() %>% 
    left_join(clean_data_table[,.(sample,library_size)]) %>% 
    filter(!is.na(library_size)) %>% 
    mutate(TPM = (NumReads )/ (library_size / 10^6)) %>% 
    unique() %>% 
  left_join(clean_data_table %>% 
    select(sample,disease_group2,tissue_clean)) %>% 
    unique() %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
      filter(Name.y %in% c("UNC13B-207",
                         "UNC13B-208",
                         "UNC13B-210")) %>%
    ggplot(aes(x = disease_group2, y = TPM, fill = Name.y)) + 
    geom_boxplot() +
    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
    facet_wrap(~tissue_clean,scales = 'free') + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) + 
    ggpubr::theme_pubr() + 
    labs(fill="",x="") +
    scale_fill_manual(values = iso_colors)

@@

df %>%
    filter(classification != "Neither") %>%
    group_by(name) %>%
    mutate(f = n/sum(n)) %>% 
  filter(grepl("FTD",name)) %>% 
  ggplot(aes(x = name, y = 100*f, fill = classification)) +
    geom_bar(stat="identity") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    xlab("") +
    ggeasy::easy_add_legend_title("") +
  scale_fill_manual(values = c("#006613ff",
                               "#ffbf00ff",
                               "#f5671cff")) +
    theme(axis.ticks.length=unit(0.1,"inch"),
        axis.line = element_line(colour = 'black', size = 1.5),
        axis.ticks = element_line(colour = "black", size = 2))

ggplot(df %>% filter(type=="Patient"), 
       aes(x = name, y = 100*fraction, fill = class2)) +
    geom_bar(stat="identity", position="dodge",color = "black") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    ggtitle("Patient RNA") +
    xlab("") +
    ggeasy::easy_add_legend_title("") +
    scale_fill_manual(values = c("#006613ff",
                               "#ffbf00ff",
                               "#f5671cff",
                               "#290340ff")) +
    theme(axis.ticks.length=unit(0.1,"inch"),
        axis.line = element_line(colour = 'black', size = 1.5),
        axis.ticks = element_line(colour = "black", size = 2))

@@

# Different guides strength

```{r}
guide_strength = fread(file.path(here::here(),"data","different_guides_cryptic_expression.csv"))
guide_strength %>% 
  melt() %>% 
  filter(`cell line` == "i5W") %>% 
   mutate(condition = fct_relevel(condition,
                              "control","tdp43_mild")) %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "value",
              fill = 'variable',
              color = 'variable',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
      scale_color_manual(
        values = c("#1C2617","#262114","#262114")
    ) +
        scale_fill_manual(
        values = c("#1576b8","#eb7f3f","#975452")
    ) +
    scale_y_continuous(labels = scales::percent) +
  ggtitle("NCRM5")

guide_strength %>% 
  melt() %>% 
  filter(`cell line` == "i11w_repeat") %>% 
   mutate(condition = fct_relevel(condition,
                              "control","tdp43_mild")) %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "value",
              fill = 'variable',
              color = 'variable',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
      scale_color_manual(
        values = c("#1C2617","#262114","#262114")
    ) +
        scale_fill_manual(
        values = c("#1576b8","#eb7f3f","#975452")
    ) +
    scale_y_continuous(labels = scales::percent) +
  ggtitle("WTC11")
```

# Dox curves and protein/RNA levels

```{r}
library(tidyverse)
library(viridis)
library(dplyr)

Protein <- read.csv(file.path(here::here(),"data","Protein_RNA_Corr_3.csv"), colClasses = c("factor","factor","numeric","factor","factor"))

Protein$Treatment <- factor(Protein$Treatment, levels=c("NT","Dox-1","Dox-2","Dox-3","Dox-4","Dox-5"))
Protein$Line_Group <- factor(Protein$Line_Group, levels=c("Protein","RNA","Correct_RNA"))

agg=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="mean")
standard_error <- function(x) sd(x) / sqrt(length(x))
agg$se=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="standard_error")$Value

limits <- aes(ymax=Value+se, ymin=Value-se)
dodge <- position_dodge(width=0.9)

agg = agg %>% 
  filter(Gene != "13A-RNA") %>% 
  mutate(Line_Group = gsub("Correct_RNA","RNA",Line_Group))

ggplot(agg, aes(y=Value, x=Treatment)) +
  geom_line(size = 2,mapping = aes(x=Treatment, y=Value,
                          linetype=Line_Group, group=Gene,color = Color_Group), size=1)+
  geom_errorbar(limits, 
                position=dodge, 
                width=0.24,
                size = 1.1)+
  xlab("") +
  ylab("") +
  ggpubr::theme_pubr()+
  scale_y_continuous(labels = scales::percent_format(scale = 1))+
  scale_colour_viridis_d(begin=0, end=1) +
  labs(linetype = "") +
  theme(axis.ticks.length=unit(0.1,"inch"),
        axis.line = element_line(colour = 'black', size = 1.5),
        axis.ticks = element_line(colour = "black", size = 2))
